---
title: "Comparison: TrialEmulation R package"
format:
  html:
    embed-resources: true
execute:
  warning: false
  message: false
---

This document runs sequential target trial emulation using the
[TrialEmulation](https://cran.r-project.org/package=TrialEmulation) R package
on the `trial_example` dataset. It is the companion to
`comparison-seqtte.qmd`, which runs the same analysis using the
`seqtte` Stata package via nbstata.

## Setup

```{r}
#| label: setup
library(TrialEmulation)
library(haven)
```

## Data

The `trial_example` dataset ships with TrialEmulation. It contains 503
individuals observed over up to 396 calendar periods.

```{r}
#| label: data
data("trial_example")
dim(trial_example)
str(trial_example)
```

**Note on eligibility.** TrialEmulation uses the `eligible` column supplied in
the data to define which person-periods serve as trial entry points. The `seqtte`
Stata package determines eligibility internally as "not previously treated"
(`A_lag != 1`). These two criteria differ slightly in `trial_example` (there are
some rows where the supplied `eligible` departs from the not-previously-treated
rule), so small differences in estimates between the two documents are expected.

## Intent-to-treat (ITT) analysis

The model adjusts for `catvarA`, `catvarB`, `nvarA`, `nvarB`, and `nvarC` at
trial entry, with quadratic polynomial terms for follow-up time and trial
(calendar) period. Standard errors are clustered by individual.

```{r}
#| label: itt
te_itt <- initiators(
  data              = trial_example,
  id                = "id",
  period            = "period",
  treatment         = "treatment",
  eligible          = "eligible",
  outcome           = "outcome",
  estimand_type     = "ITT",
  outcome_cov       = ~ catvarA + catvarB + nvarA + nvarB + nvarC,
  model_var         = "assigned_treatment",
  save_weight_models = FALSE
)
```

### Regression table

```{r}
#| label: itt-table
print(te_itt)
```

### Treatment effect

```{r}
#| label: itt-coef
m <- te_itt$model
coef_trt   <- coef(m)["assigned_treatment"]
se_trt     <- sqrt(vcov(m)["assigned_treatment", "assigned_treatment"])
or_trt     <- exp(coef_trt)
ci_trt     <- exp(coef_trt + c(-1, 1) * qnorm(0.975) * se_trt)

cat(sprintf(
  "assigned_treatment log-OR: %.4f (SE %.4f)\nOR: %.4f (95%% CI %.4f to %.4f)\n",
  coef_trt, se_trt, or_trt, ci_trt[1], ci_trt[2]
))
```

## Save data for Stata comparison

The dataset is saved as a Stata `.dta` file so it can be loaded in
`comparison-seqtte.qmd`.

```{r}
#| label: save-dta
write_dta(trial_example, "trial_example.dta")
cat("Saved trial_example.dta\n")
```
